Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface

### 2022-03-09 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Clean up text. Add figure captions

@@

Most interfaces to gas-accepting sputter ion sources are a closed inlet design. A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the sample volume and the length and diameter of the capillary. Flow may be adjusted by changing the pressure by varying the sample volume.

We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere. This defines the delivery pressure as 1 ATM and therefore flow is controlled by the capillary and is not adjustable. The flow of CO2 to the source may be varied by controlling the ratio of CO2 to helium in the sample stream. An open inlet has the advantage of simplicity; flow of CO2 to the source is preset and there are no separate sample transfer and dilution steps. This may translate to greater throughput by eliminating the time required for sample transfer, dilution, and setting of delivery pressure. A disadvantage of the open split is that the flow of CO2 to the source varies with sample size (CO2:He) and changes over the course of a measurement due to dilution of the sample by displacement helium. Our method compensates for this by limiting CO2 flow to a range that produces relatively constant C- ion current and stable ratios. Salazar and Szidat, Roberts gasbench, other open split methods.

@@

The open split interface supplies a helium-CO2 mixture to a split where a capillary delivers gas to the source and the rest is vented. A flow of helium is used to displace sample CO2 to the delivery capillary in the split. Our open-split interface is constructed of simple, off-the-shelf parts. The core of the design is a dual needle and split housed in a VICI cross. Displacement helium flows in through one side of the cross, into the sample vial through 1/16" stainless steel tubing, which forms the outer flow path of the concentric dual needle. The mixture of sample CO2 and displacement helium flows up a needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle. The split is made from a VICI Luer adapter fitted to the outflow of the dual needle with the source delivery capillary inserted (Figure 1). The split has fairly low volume (< 1 mL). Testing is needed to determine minimum sample flow to avoid atmosphere entering the split. TODO: move flow ratio test to results. The system is currently not automated, but may be easily mounted on the arm of a gasbench-style autosampler.

@@

Figure: Current vs flow

@@

Figure: current dependent blank

@@

Figure: justification for large blank

@@

Figure: sample size vs current vs precision and accuracy

@@

Two more recent carbonate tests produced stable currents and backgrounds. Since sample, target blank, and sample blank currents were relatively constant across samples, a large blank correction can be used to correct results. Samples from both runs showed a linear relationship between difference from expected Fm and Fm of the sample, but in one sample set the relationship was positive, and in the other, negative. Since the intercept with 0-offset was close to 0 Fm, I think this is a problem with normalization, rather than blank correction. Standards run at the end of the day are higher than those run at the beginning, which may give clues as to what's going on.

@@

figure: difference plot

@@

Figure: bland-altman of HGIS vs graphite.

@@

Figure: age model comparison graphite HGIS

@@

>>>>

Most interfaces to gas-accepting sputter ion sources are a closed inlet design. A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the sample volume and the length and diameter of the capillary. Flow may be adjusted by changing the pressure by varying the sample volume. We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere. This defines the delivery pressure as 1 ATM and therefore flow is controlled by the capillary and is not adjustable. The flow of CO2 to the source may be varied by controlling the ratio of CO2 to helium in the sample stream. An open inlet has the advantage of simplicity; flow of CO2 to the source is preset and there are no separate sample transfer and dilution steps. This may translate to greater throughput by eliminating the time required for sample transfer, dilution, and setting of delivery pressure. A disadvantage of the open split is that the flow of CO2 to the source varies with sample size (CO2:He) and changes over the course of a measurement due to dilution of the sample by displacement helium. Our method compensates for this by limiting CO2 flow to a range that produces relatively constant C- ion current and stable ratios. Salazar and Szidat, Roberts gasbench, other open split methods.

@@

@@

The open split interface supplies a helium-CO2 mixture to a split where a capillary delivers gas to the source and the rest is vented. A flow of helium is used to displace sample CO2 to the delivery capillary in the split. Our open-split interface is constructed of simple, off-the-shelf parts. The core of the design is a dual needle and split housed in a VICI cross. Displacement helium flows in through one side of the cross, into the sample vial through 1/16" stainless steel tubing, which forms the outer flow path of the concentric dual needle. The mixture of sample CO2 and displacement helium flows up a needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle. The split is made from a VICI Luer adapter fitted to the outflow of the dual needle with the source delivery capillary inserted (Figure 1). The split has fairly low volume (< 1 mL). Testing is needed to determine minimum sample flow to avoid atmosphere entering the split. TODO: move flow ratio test to results. The system is currently operated manually, but may be easily mounted on the arm of a gasbench-style autosampler to enable automated operation.

@@

Figure1: Current vs flow

@@

Figure2: current dependent blank

@@

Figure3: justification for large blank

@@

Figure4: sample size vs current vs precision and accuracy

@@

Two more recent carbonate tests produced stable currents and backgrounds. Sample, target blank, and sample blank currents were relatively constant across samples, so a large blank correction can be used to correct results. Samples from both runs showed a linear relationship between difference from expected Fm and Fm of the sample, but in one sample set the relationship was positive, and in the other, negative. Since the intercept with 0-offset was close to 0 Fm, I think this is a problem with normalization, rather than blank correction. Standards run at the end of the day are higher than those run at the beginning, which may give clues as to what's going on.

@@

Table: summary of reference materials

figure5: difference plot

@@

Figure6: bland-altman of HGIS vs graphite.

@@

Figure7: age model comparison graphite HGIS

@@

# Figure captions

Fig. 1. Performance of the NOSAMS MCSNICS hybrid gas ion source.

Fig. 2. Assessment of blanks. (a) Current dependent blank. (b) Fraction modern of large, radiocarbon dead samples compared to sample current. Blank rises with decreasing current, as expected, but may be treated as a single blank with a simple correction for full-size samples producing greater than XX uA C- current.

Fig. 3. Relationship of sample size to C- ion current. There is no strong relationship between size and current for samples larger than XX ug C.

Fig 4. Agreement of standards with graphite consensus values.

Fig 5. Comparison of Haiti core samples measured as graphite.
